---
title: "Model Feeding Time Present <> Absent"
author: "Andreas Eich"
format: html
editor: visual
---


### Libraries

```{r}
library(here) # easier paths, here() function starts in the folder of the RStudio project
library(tidyverse)
library(brms)
library(tidybayes)
library(ggtern)
library(kableExtra)
library(janitor)
library(bayesplot)
library(emmeans)
library(marginaleffects)
library(htmlTable)

# get custom functions
source(here("scripts/functions.R"))
```

### Get data

```{r}
dat_sec_feeding <- readRDS(here("data/clean/clean_data.rds"))
```

### Raw Data

Mean feeding time **per video** ± SE

```{r}
dat_sec_feeding %>%
  group_by(size_class, shark_in_video, year, video) %>%
  #mean feeding time per video
    summarise(mean_sec_feeding_video = mean(sec_feeding, na.rm = T)) %>%
   group_by(size_class, shark_in_video, year) %>%
  #mean feeding time per group
  summarise(
    mean_sec_feeding = mean(mean_sec_feeding_video, na.rm = T),
    sd = sd(mean_sec_feeding_video, na.rm = T),
    n = n(),
    se = ifelse(n >= 3, sd / sqrt(n), NA)
  ) %>%
  ggplot() +
  geom_errorbar(aes(
    x = shark_in_video,
    ymin = mean_sec_feeding - se,
    ymax = mean_sec_feeding + se,
    col = factor(shark_in_video),
    width = .4
  )) +
  geom_point(aes(
    x = shark_in_video,
    y = mean_sec_feeding,
    col = factor(shark_in_video)
  )) +
  facet_grid(year ~ size_class , scales = "free") +
  theme_clean()
```

### Models Small & Medium

```{r}
dat_sec_feeding_SM <- dat_sec_feeding %>%
  filter(size_class != "Large") %>% 
  droplevels()


fit_sec_feeding_SM <- brm(
  bf(
    sec_feeding ~ year * shark_in_video * size_class +
      s(log_n_frame_total) + 
                  habitat + 
                  s(log_depth) +
                  (1|video), 
    hu ~ year * shark_in_video
  ),
             control = list(max_treedepth = 14, 
                          adapt_delta = 0.9),
  family = hurdle_gamma(link = "log"),
  data = dat_sec_feeding_SM,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = 4,
  seed = 123,
  file_refit = "on_change",
  threads = threading(2, static = T),
  file = here("models/fit_sec_feeding_SM"),
  backend = "cmdstanr"
)
```

Assess Model

```{r}
summary(fit_sec_feeding_SM)#ok 127 divergent transitions, Rhat ok

bayesplot::ppc_dens_overlay(y = log1p(dat_sec_feeding_SM$sec_feeding), 
                            yrep = log1p(posterior_predict(fit_sec_feeding_SM)[1:10,]))

bayes_R2(fit_sec_feeding_SM)#0.1207585
```

#### Plot Results for year shark_in_video size_class

Emmeans does not include zero inflation(?) \> manually

```{r}

epred_sec_feeding_SM <-  epred_draws(
  fit_sec_feeding_SM,
  newdata = expand_grid(
    year = dat_sec_feeding_SM$year %>% unique(),
    shark_in_video = dat_sec_feeding_SM$shark_in_video %>% unique(),
    size_class = c("Small", "Medium"),
    habitat = dat_sec_feeding_SM$habitat %>% unique(),
    log_depth = mean(dat_sec_feeding_SM$depth) %>%
      log(),
    log_n_frame_total = mean(dat_sec_feeding_SM$n_frame_total) %>%
      log()
  ),
  re_formula = NA
) %>%
  #average over habitat
  group_by(year,
           shark_in_video,
           size_class,
           log_depth,
           log_n_frame_total,
           .draw) %>%
  summarise(.epred = mean(.epred))

```

Plot

```{r}

plot_sec_feeding_SM <- ggplot(data = epred_sec_feeding_SM, 
                                 aes(x = year, 
                                     col = paste(shark_in_video, size_class, sep = "_"))) +
  geom_point(
    data = dat_sec_feeding_SM,
    aes(y = sec_feeding),
    position =  position_jitterdodge(jitter.width = .2, dodge.width = .7),
    alpha = .1,
    size = 3,
    stroke = 0
  ) +
  stat_pointinterval(
    aes(y = .epred),
    point_interval = mode_hdi,
    .width = c(.95, .89),
    position = position_dodge(width = .7),
    show.legend = F
  ) +
  scale_colour_manual(values = cols_sharks,
                      labels = c("absent", "absent", "present", "present")) +
  guides(color = guide_legend(override.aes = list(alpha = 1), title = "Shark")) +
  labs(x = NULL, y = "Feeding time (sec)") +
  facet_grid( ~ factor(size_class, levels = c("Small", "Medium")), scales = "free") +
  scale_y_continuous(trans = 'log10') +
  theme_clean(); plot_sec_feeding_SM
```

And save

```{r}
ggsave(filename = "sec_feeding_SM.pdf",
       plot = plot_sec_feeding_SM, 
       path = here("plots"),
       width = 15, height = 9, units = "cm")
```

##### Conditional effects for shark_in_video per year and size_class

```{r}
epred_sec_feeding_SM %>% 
  pivot_wider(names_from = "shark_in_video", values_from = ".epred") %>% 
  mutate(contrast = "absent - present",
         .value = absent - present) %>% 
  group_by(contrast, size_class, year) %>%
  # median_hdi returns wrong groups? => manually
  summarise(median = median(.value),
            bayestestR::hdi(.value)) %>% 
  rename(".lower"  = "CI_low",
         ".upper"  = "CI_high") %>% 
  select(-CI) %>%
  mutate(important = ifelse(.upper > 0 & .lower > 0 |
                       .upper < 0 & .lower < 0, "*","")) %>% 
  mutate_if(is.numeric, txtRound,2) %>% 
  arrange(desc(size_class)) %>% 
  htmlTable(rnames = FALSE)

```

No effect

##### Conditional effects for habitat

```{r}
epred_sec_feeding_SM_habitat <- epred_draws(
  fit_sec_feeding_SM,
  newdata = expand_grid(
    year = dat_sec_feeding_SM$year %>% unique(),
    shark_in_video = dat_sec_feeding_SM$shark_in_video %>% unique(),
    size_class = c("Small", "Medium"),
    habitat = dat_sec_feeding_SM$habitat %>% unique(),
    log_depth = mean(dat_sec_feeding_SM$depth) %>%
      log(),
    log_n_frame_total = mean(dat_sec_feeding_SM$n_frame_total) %>%
      log()
  ),
  re_formula = NA
) %>%
  #average over habitat
  group_by(habitat,
           .draw) %>%
  summarise(.epred = mean(.epred)) %>%
  pivot_wider(names_from = "habitat", values_from = ".epred") %>% 
  mutate(contrast = "near reef - reef",
         .value = `near reef` - reef) 

#Table
epred_sec_feeding_SM_habitat %>% 
  group_by(contrast) %>%
  median_hdi(.value) %>% 
  mutate(important = ifelse(.upper > 0 & .lower > 0 |
                       .upper < 0 & .lower < 0, "*","")) %>% 
  select(-.width, -.point, -.interval) %>% 
  mutate_if(is.numeric, txtRound,2) %>% 
  htmlTable(rnames = FALSE)

#Plot
epred_sec_feeding_SM_habitat %>% 
  ggplot()+
  geom_hline(yintercept = 0, linetype = "11")+
  stat_pointinterval(
    aes(y = .value, x = 1),
    point_interval = mode_hdi,
    .width = c(.95, .89),
    position = position_dodge(width = .7),
    show.legend = F
  )+
  theme_clean()

```

No effect

##### Conditional effects for n fish in frame

```{r}
epred_sec_feeding_SM_ntotal <- epred_draws(
  fit_sec_feeding_SM,
  newdata = expand_grid(
    year = dat_sec_feeding_SM$year %>% unique(),
    shark_in_video = dat_sec_feeding_SM$shark_in_video %>% unique(),
    size_class = c("Small", "Medium"),
    habitat = dat_sec_feeding_SM$habitat %>% unique(),
    log_depth = mean(dat_sec_feeding_SM$depth) %>%
      log(),
    log_n_frame_total = seq(min(dat_sec_feeding_SM$n_frame_total),
                            max(dat_sec_feeding_SM$n_frame_total),
                            length = 100) %>%
      log()
  ),
  re_formula = NA
) %>%
  #average over habitat
  group_by(log_n_frame_total,
           .draw) %>%
  summarise(.epred = mean(.epred))


plot_sec_feeding_SM_ntotal <- ggplot(data = dat_sec_feeding_SM, 
                                     aes(x = log_n_frame_total))+
  stat_lineribbon(data = epred_sec_feeding_SM_ntotal,
                  aes(y = .epred),
                  .width = c(.95),
                  fill = "lightblue", show.legend = F)+
  geom_point(aes(y = sec_feeding), alpha = .1)+
  labs(x = "log(n fish in frame)", y = "Feeding time (sec)")+
  theme_clean()+
  theme(legend.position = "none")

plot_sec_feeding_SM_ntotal

ggsave(filename = "sec_feeding_SM_ntotal.pdf",
       plot = plot_sec_feeding_SM_ntotal, 
       path = here("plots/SI"),
       width = 20, height = 15, units = "cm")
```

##### Conditional effects for depth

```{r}

log_n_frame_total = seq(min(dat_sec_feeding_SM$n_frame_total),
                            max(dat_sec_feeding_SM$n_frame_total),
                            length = 100) %>%
      log()


epred_sec_feeding_SM_depth <- epred_draws(
  fit_sec_feeding_SM,
  newdata = expand_grid(
    year = dat_sec_feeding_SM$year %>% unique(),
    shark_in_video = dat_sec_feeding_SM$shark_in_video %>% unique(),
    size_class = c("Small", "Medium"),
    habitat = dat_sec_feeding_SM$habitat %>% unique(),
    log_n_frame_total = mean(dat_sec_feeding_SM$n_frame_total) %>%
      log(),
    log_depth = seq(min(dat_sec_feeding_SM$depth),
                            max(dat_sec_feeding_SM$depth),
                            length = 100) %>%
      log()
  ),
  re_formula = NA
) %>%
  group_by(log_depth,
           .draw) %>%
  summarise(.epred = mean(.epred))


plot_sec_feeding_SM_depth <- ggplot(data = dat_sec_feeding_SM, 
                                     aes(x = log_depth))+
  stat_lineribbon(data = epred_sec_feeding_SM_depth,
                  aes(y = .epred),
                  .width = c(.95),
                  fill = "lightblue", show.legend = F)+
  geom_point(aes(y = sec_feeding), alpha = .1)+
  labs(x = "log(depth)", y = "Feeding time (sec)")+
  theme_clean()+
  theme(legend.position = "none")

plot_sec_feeding_SM_depth

ggsave(filename = "sec_feeding_SM_depth.pdf",
       plot = plot_sec_feeding_SM_depth, 
       path = here("plots/SI"),
       width = 20, height = 15, units = "cm")
```
No impact

```{r}
plot_epred_sec_feeding_SM_ntotal <- plot_predictions(
  fit_sec_feeding_SM,
  condition = c("log_n_frame_total"),
  conf_level = .95
) +
  geom_point(data = dat_sec_feeding_SM, aes(x = log_n_frame_total, y = sec_feeding), alpha = .1) +
  labs(x = "log(n fish in frame)", y = "Feeding time (sec)") +
  theme_clean()


plot_epred_sec_feeding_SM_ntotal

ggsave(filename = "sec_feeding_SM_ntotal.pdf",
       plot = plot_epred_sec_feeding_SM_ntotal, 
       path = here("plots/SI"),
       width = 20, height = 15, units = "cm")
```

No impact, mostly 0
