---
title: "Model Bout Rate Present <> Absent"
author: "Andreas Eich"
format: html
editor: visual
---

### Libraries

```{r}
library(here) # easier paths, here() function starts in the folder of the RStudio project
library(tidyverse)
library(brms)
library(tidybayes)
library(ggtern)
library(kableExtra)
library(janitor)
library(bayesplot)
library(emmeans)
library(marginaleffects)
library(htmlTable)

# get custom functions
source(here("scripts/functions.R"))
```

### Get data

Sometimes bouts/bites \> 1 when feeding time = 0, sometimes feeding time \> 0 but bout/bites = 0

```{r}
dat_boutrate <- readRDS(here("data/clean/clean_data.rds")) %>% 
  filter(bite > 0 & sec_feeding > 0)
```

Calculate bouts per sec

```{r}
dat_boutrate <- dat_boutrate %>% 
  mutate(boutrate = bout/sec_feeding)
```

### Raw Data

Mean feeding time **per video** ± SE

```{r}
dat_boutrate %>%
  group_by(size_class, shark_in_video, year, video) %>%
  #mean feeding time per video
    summarise(mean_boutrate_video = mean(boutrate, na.rm = T)) %>%
   group_by(size_class, shark_in_video, year) %>%
  #mean feeding time per group
  summarise(
    mean_boutrate = mean(mean_boutrate_video, na.rm = T),
    sd = sd(mean_boutrate_video, na.rm = T),
    n = n(),
    se = ifelse(n >= 3, sd / sqrt(n), NA)
  ) %>%
  ggplot() +
  geom_errorbar(aes(
    x = shark_in_video,
    ymin = mean_boutrate - se,
    ymax = mean_boutrate + se,
    col = factor(shark_in_video),
    width = .4
  )) +
  geom_point(aes(
    x = shark_in_video,
    y = mean_boutrate,
    col = factor(shark_in_video)
  )) +
  facet_grid(year ~ size_class , scales = "free") +
  theme_clean()
```

Only possible for small fish

### Models Small

Few observations, focus on year \* shark_in_video

```{r}
dat_boutrate_S <- dat_boutrate %>%
  filter(size_class == "Small") %>% 
  select(-size_class)


fit_boutrate_S <- brm(
    boutrate ~ year * shark_in_video +
                  (1|video),
             control = list(max_treedepth = 14, 
                          adapt_delta = 0.99),
  family = Gamma(link = "log"),
  data = dat_boutrate_S,
  iter = 4000,
  warmup = 2000,
  chains = 4,
  cores = 4,
  seed = 123,
  file_refit = "on_change",
  threads = threading(2, static = T),
  file = here("models/fit_boutrate_S"),
  backend = "cmdstanr"
)
```

Assess Model

```{r}
summary(fit_boutrate_S)#Rhat ok

bayesplot::ppc_dens_overlay(y = log1p(dat_boutrate_S$sec_feeding), 
                            yrep = log1p(posterior_predict(fit_boutrate_S)[1:10,]))

bayes_R2(fit_boutrate_S)#0.476964
```

#### Plot Results for year shark_in_video

Get Emmeans values

```{r}
epred_boutrate_S <- emmeans(fit_boutrate_S,
                              ~year * shark_in_video, 
                              typ = "response") %>%   
  gather_emmeans_draws() %>% 
  mutate(.value = exp(.value))# gather_emmeans_draws() removes back-transformation
```

Plot

```{r}
plot_boutrate_S <- ggplot(data = epred_boutrate_S %>% 
      mutate(size_class = "Small"), 
                                 aes(x = year, 
                                     col = shark_in_video)) +
  geom_point(
    data = dat_boutrate_S %>% 
      mutate(size_class = "Small"),
    aes(y = boutrate),
    position =  position_jitterdodge(jitter.width = .2, dodge.width = .7),
    alpha = .1,
    size = 3,
    stroke = 0
  ) +
  stat_pointinterval(
    aes(y = .value),
    point_interval = mode_hdi,
    .width = c(.95, .89),
    position = position_dodge(width = .7),
    show.legend = F
  ) +
  scale_colour_manual(values = cols_sharks_S,
                      labels = c("absent", "present")) +
  guides(color = guide_legend(override.aes = list(alpha = 1), title = "Shark")) +
  labs(x = NULL, y = "Bouts per second") +
 # scale_y_continuous(trans = 'log10') +
  facet_grid( ~ size_class, scales = "free") +
  theme_clean(); plot_boutrate_S
```

And save

```{r}
ggsave(filename = "boutrate_S.pdf",
       plot = plot_boutrate_S, 
       path = here("plots"),
       width = 15/1.7, height = 9, units = "cm")
```

#### Conditional effects for shark_in_video per year

```{r}
emmeans(fit_boutrate_S, ~ shark_in_video |  year, type = "response") %>%
  regrid() %>%
  pairs() %>%
  gather_emmeans_draws() %>%
  group_by(contrast, year) %>%
    summarise(median = median(.value),
            bayestestR::hdi(.value,ci = 0.95)) %>% 
  rename(".lower"  = "CI_low",
         ".upper"  = "CI_high") %>% 
    select(-CI) %>%
  mutate(important = ifelse(.upper > 0 & .lower > 0 |
                       .upper < 0 & .lower < 0, "*","")) %>% 
  mutate_if(is.numeric, txtRound,2) %>% 
  htmlTable(rnames = FALSE)

```

```{r}
emmeans(fit_boutrate_S, ~ shark_in_video |  year, type = "response") %>%
  regrid() %>%
  pairs() %>%
  gather_emmeans_draws() %>%
  filter(.value > 0) %>% 
  group_by(contrast, year) %>% 
  summarise(above_0 = n()/8000) %>% 
  mutate_if(is.numeric, txtRound,2) %>% 
  htmlTable(rnames = FALSE)
```

